---
title: "Final_marginal_likelihoods.R"
author: "Eric Crandall"
date: "11/6/2016"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# a function for calculating model selection statistics
bfcalcs<-function(df,ml="bezier.corrected"){
  df$thermodynamic<-as.numeric(df$thermodynamic)
  df$bezier.corrected<-as.numeric(df$bezier.corrected)
  df$harmonic<-as.numeric(df$harmonic)
    mlcol<-df[,ml] 
	bmvalue<-mlcol[which.max(mlcol)]
	lbf<-2*(mlcol-bmvalue)
	choice<-rank(-mlcol)
	modelprob<-exp(lbf/2)/sum(exp(lbf/2))
	dfall<-cbind(df,lbf,choice,modelprob)
	return(dfall)
}	
```

## Harvesting likelihoods

Here I am going through the final three runs (rounds 6,7,8) and harvesting out the likelihoods into a likelists object.

```{r harvest_likelihoods}
wd<-"/Users/eric/Datasets/ToBo_Migrate_runs/"
setwd(wd)



likelists<-list()
for(r in 6:8){
  wd1<-paste(wd,"round",r,sep="")
  setwd(wd1)
  # start here if you just want to do one list
  likelist<-list() #initialize an empty list
  for(f in list.files()){
    wd2<-file.path(wd1,f)
    marglike<-data.frame(model=character(0),thermodynamic=numeric(0),bezier.corrected=numeric(0),harmonic.mean=numeric(0),stringsAsFactors=F) #initialize a data frame to take the values
    l=1 #initialize l
    for(i in c("stepping.stone","stepping.stone.1param","2currents","1current","hi_lo","n.island","empirical","panmixia")){ #i<-"stepping.stone"
      wd3<-file.path(wd2,i)
      print(wd3)
      if(!file.exists(wd3)){next}
      setwd(wd3)
      outfile<-scan(file=paste(f,"outfile.txt",sep="_"),what="character",sep="\n") #scan in the outfile, separating at each newline
      
        if(grepl("gym",f)==T){
        all.locus.line<-outfile[grep("\\[Scaling",outfile,value=F)-1] #find the line with all three values on it (it comes right before the line that has the scaling factor on it)
        all.locus.line<-strsplit(all.locus.line," +")
        if(length(all.locus.line)==0){next}
        thermo<-as.numeric(all.locus.line[[1]][3])
        bezier<-as.numeric(all.locus.line[[1]][4])
        harmo<-as.numeric(all.locus.line[[1]][5])
        marglike[l,]<-c(i,thermo,bezier,harmo) #add this as a row to the data frame
        l=l+1
        likelist[[f]]<-marglike
        next
      }
      
      #get the result from thermodynamic integration
      thermoline<-grep("(1a)",outfile,value=T) #find the line with the thermodynamic likelihood on it
      if(length(thermoline)==0){next}
      thermoline<-strsplit(thermoline,split="=",fixed=T) #split it up
      thermo<-as.numeric(substr(thermoline[[1]][2],start=1,stop=12)) #grab the thermodynamic likelihood
      bezistart<-grep("\\(",strsplit(thermoline[[1]][2],split="")[[1]])+1
      bezier<-as.numeric(substr(thermoline[[1]][2],start=bezistart,stop=bezistart+11)) #and the bezier-corrected value
      #get the harmonic mean
      harmoline<-grep("\\(2\\) H",outfile,value=T) #find the line with harmonic mean likelihood on it
      harmoline<-strsplit(harmoline,split="=",fixed=T) #grab the harmonic mean
      harmo<-as.numeric(substr(harmoline[[1]][2],start=1,stop=12))
      marglike[l,]<-c(i,thermo,bezier,harmo) #add this as a row to the data frame
      l=l+1
    }
    

    
    likelist[[f]]<-marglike #add the dataframe to the list
  }
  
# stop here if you just want one round  
  likelists[[r]]<-likelist
}
setwd(wd)
```

## Model Selection
On each round, and then bind them into a list
```{r model_selection}
modeltable1<-lapply(likelists[[6]],bfcalcs)
modeltable2<-lapply(likelists[[7]],bfcalcs)
modeltable3<-lapply(likelists[[8]],bfcalcs)
modeltables<-list("round1"=modeltable1,"round2"=modeltable2,"round3"=modeltable3)

head(modeltables[[1]],1)

#Remove anon9 from the dataset
modeltables$round1$anon9_CR_blank<-NULL
modeltables$round2$anon9_CR_blank<-NULL
modeltables$round3$anon9_CR_blank<-NULL
```


## Plot marginal likelihood results from each of 3 runs
```{r}
#plot model selection graphics (blue dots with lines) - lines denote confidence intervals
m<-c("panmixia","hi_lo","1current","2currents","n.island","structure_panmixia","structure_currents","stepping.stone.1param","stepping.stone","empirical")
m<-c("panmixia","hi_lo","1current","2currents","n.island","stepping.stone.1param","stepping.stone","empirical")


pdf(file = "thermodynamic_marginal_likelihoods_final.pdf",width=8.5,height=3)
means<-list()
#plots<-list()
for(dataset in names(modeltables[[1]])){
  likes<-rbind(cbind(modeltables[[1]][[dataset]],rep=1), cbind(modeltables[[2]][[dataset]],rep=2), cbind(modeltables[[3]][[dataset]], rep=3))
  likes$model<-factor(likes$model, m)
  likes<-likes[!(is.na(likes$model)),]
  
  #likes<-likes[which(likes$bezier.corrected > max(likes$bezier.corrected)-100),]
  y.mean<-as.vector(by(likes$bezier.corrected,likes$model,mean))
  y.sd<-as.vector(by(likes$bezier.corrected,likes$model,sd))
  y.min<-y.mean-((y.sd/sqrt(3))*4.303)
  y.max<-y.mean+((y.sd/sqrt(3))*4.303)

    
  likes.mean<-data.frame(model=factor(m,m),y.mean,y.min,y.max,y.sd)
  means[[dataset]]<-likes.mean
  
  #l<-ggplot(data=likes, aes(x=model,y=bezier.corrected,colour=factor(rep), 
   #                                 shape=factor(rep), size=20 ))
  l<-ggplot(data=likes, aes(x=model,y=bezier.corrected))
  
  l<-l+geom_point(colour="blue", size=3)+
    geom_pointrange(data=likes.mean,y=y.mean,ymin=y.min,ymax=y.max, size=0.5)+
    scale_x_discrete(drop=FALSE)+
    theme(axis.text.y = element_text(size=16),legend.position="none",axis.title.x=element_text(size=16),axis.title.y=element_blank(),plot.title=element_text(size=20))+ggtitle(dataset)+ylab("Marginal Log-Likelihood")+
    coord_fixed(0.1)+ coord_flip()
  print(l)
#  plots<-c(plots,l)
}
dev.off()
```


## Take the mean marginal likelihoods over three runs and calculate model selection tables.
```{r}
library(perm)

meanlikes<-list()
permtests<-data.frame(matrix(ncol = 13, nrow = 0))
colnames(permtests)<-c(m,"k-means p-value","Best Mean","Second Best Mean","p-value Best Mean > Second Mean", "Log Bayes Factor")
for(dataset in names(modeltables[[1]])){
  likes<-rbind(cbind(modeltables[[1]][[dataset]],rep=1), cbind(modeltables[[2]][[dataset]],rep=2), cbind(modeltables[[3]][[dataset]], rep=3))
  likes$model<-factor(likes$model, m)
  likes<-likes[!(is.na(likes$model)),]
  likes$model<-factor(likes$model, m)
  y.mean<-as.vector(by(likes$bezier.corrected,likes$model,mean)) #take the means for each model
  kmeans.p<-permKS(x = likes$bezier.corrected, g = likes$model, method="exact.mc")$p.value #k-means non-parametric test to compare all means
  bestmean<-m[order(y.mean,decreasing=T)[1]] # get the highest mean
  secondmean<-m[order(y.mean,decreasing=T)[2]] #get the second best mean
  ttest<-permTS(x=likes$bezier.corrected[which(likes$model==bestmean)], y=likes$bezier.corrected[which(likes$model==secondmean)], alternative="greater", method="exact.mc")
  lbf<-2*ttest$estimate
  
  values<-rbind(y.mean)
  values<-cbind(values,kmeans.p,bestmean,secondmean,ttest$p.value,lbf)
  
  permtests[dataset,]<-values

	bmvalue<-y.mean[which.max(y.mean)]
	lbf<-2*(y.mean-bmvalue)
	modelrank<-rank(-y.mean)
	modelprob<-exp(lbf/2)/sum(exp(lbf/2),na.rm=T)
	dfall<-cbind("Mean Bezier Corrected ML"=as.numeric(y.mean),"LBF"=as.numeric(lbf),"Rank"=as.numeric(modelrank),"Model_Probability"=as.numeric(modelprob))
	row.names(dfall)<-m
	dfall<-as.data.frame(dfall)
	meanlikes[[dataset]]<-dfall
  
}
```

## Calculate the best models table and save it
```{r}
getmodels2<-function(dfr){
  model1<-row.names(dfr)[which(dfr$Rank==1)]
  model2<-row.names(dfr)[which(dfr$Rank==2)]
  modelprob1<-dfr$Model_Probability[which(dfr$Rank==1)]
  modelprob2<-dfr$Model_Probability[which(dfr$Rank==2)]
  print(model1)
  print(model2)
  print(modelprob1)
  c(model1,modelprob1,model2,modelprob2)
}

bestmodeltable<-lapply(meanlikes,getmodels2)
bestmodeltable<-t(as.data.frame(bestmodeltable))
colnames(bestmodeltable)<-c("bestmodel","bestmodelprob","secondbestmodel","secondbestmodelprob")
bestmodeltable<-cbind(bestmodeltable,permtests[9],permtests[12])


write.csv(bestmodeltable,file="Final_BestModeltable_Using_Means_of_3reps.csv")
write.csv(permtests,file="Permutation_ttest_results.csv")
save(meanlikes,file="Final_Modeltable_Using_Means_of_3reps.R")
```


## Plot Model Probability

```{r model_probability}
m<-c("stepping.stone","stepping.stone.1param","2currents","1current","hi_lo","empirical","n.island","panmixia")

#read in the t-test results
tt<-read.table(file="~/Dropbox/Crandall_tobo/scripts/Final_BestModeltable_Using_Means_of_3reps.csv",header=T,sep=",")

means<-data.frame(dataset=character(),species=character(),common=character(),models=character(),modelprob=numeric())
names<-read.csv("Species_names.csv", stringsAsFactors = F)

for(dataset in names(modeltables[[1]])){
  #make a long dataset of model probabilities from each dataset
  species<-names[which(names$Working==dataset),2] #lookup the species name
  common<-names[which(names$Working==dataset),3] #lookup the common name
  likes<-rbind(cbind(modeltables[[1]][[dataset]],rep=1), cbind(modeltables[[2]][[dataset]],rep=2), cbind(modeltables[[3]][[dataset]], rep=3))
  likes$model<-factor(likes$model, m)
  likes<-likes[!(is.na(likes$model)),]
  #take the mean model probability
  y.mean<-as.vector(by(likes$modelprob,likes$model,mean))
  #put this back in a data frame with model
  likes.mean<-data.frame(dataset=dataset,species=species,common=common,model=factor(m,m),modelprob=y.mean)
  means<-rbind(means,likes.mean)
}

#CODE FOR SORTING SPECIES
#take the value for stepping stone from each model and sort on it
#species<-unique(as.character(means$species)) #make a character vector of all species

#set mean probs for stepping stone that are less than 0.01 to 0 so I can also sort on panmixia
#means[means$model=="stepping.stone" & means$modelprob<0.01,"modelprob"]<-0
#sort it by value of stepping.stone
#species2<-species[order(-tt$ttest_pvalue, #means[means$model=="stepping.stone","modelprob"], #means[means$model=="stepping.stone.1param","modelprob"],
#means[means$model=="hi_lo","modelprob"],
#means[means$model=="empirical","modelprob"],
#means[means$model=="n.island","modelprob"],
#                        decreasing=T)] 
#
#means$model<-factor(means$model,levels=m,labels=c("k <= 14, Stepping-Stone, ~40 parameter","k <= 14, Stepping-Stone, 2 parameter","k = 3, 2 Currents, 7 parameter","k = 2, 1 Current, 4 parameter", "k = 2, High/Low Islands, 4 parameter","k = ? Empirical Structure, ? parameter","k <= 14, n-Island, 2 parameter","k=1, Panmixia, 1 parameter"))

# This is the order I settled on - significant first.
species2<-c("Cellana talcosa", "Ophiocoma pica", "Chaetodon miliaris", 
"Zebrasoma flavescens", "Caranx melampygus", "Heterocentrotus mammillatus", 
"Mulloidichthys vanicolensis", "Panulirus penicillatus", "Calcinus hazletti", 
"Chaetodon multicinctus", "Chaetodon lunulatus", "Holothuria atra", 
"Cellana exarata", "Holothuria whitmaei", "Halichoeres ornatissimus", "Acanthurus olivaceus","Gymnothorax flavimarginatus","Squalus mitsukurii", 
"Abudefduf abdominalis", "Mulloidichthys flavolineatus", "Stenella longirostris", "Panulirus marginatus",
"Chaetodon fremblii", "Calcinus seurati", "Epinephelus quernus", "Lutjanus kasmira", 
"Acanthurus nigroris",  "Etelis marshi", "Parupeneus multifasciatus", "Ctenochaetus strigosus","Triaenodon obesus", "Abudefduf vaigiensis", "Stegastes fasciolatus", 
"Pristipomoides filamentosus", "Cellana sandwicensis", "Dascyllus albisella", 
"Acanthurus nigrofuscus", "Acanthaster planci", "Ophiocoma erinaceus", 
"Gymnothorax undulatus", "Myripristis berndtii")

means$species<-factor(means$species,levels=species2)
means$model<-factor(means$model,levels=m,labels=c("Stepping-Stone: Full Model","Stepping-Stone: 2 parameter","Regional: 2 Currents","Regional: 1 Current", "Regional: High/Low Islands","Regional: Empirical Structure","n-Island","Panmixia"))

library(ggplot2)
library(RColorBrewer)

my.cols<-brewer.pal(8,"RdYlBu")
my.cols[8]<-"#7f7f7f"
my.cols[5]<-"#FFFFBF"
my.cols[6]<-"#FFFF00"


mprobs<-ggplot(data=means,mapping=aes(x=species,y=modelprob,fill=model)) + 
  geom_bar(stat="identity") + scale_fill_manual(values=my.cols, guide = guide_legend(title = "Migrate Models")) + 
  theme(axis.text.x=element_text(angle=90, hjust=1,vjust=0.5, color=c(rep("grey25",26),rep("grey70",15))), legend.text = element_text(size = 14), legend.title = element_text(size = 20), axis.text.y= element_text(size = 14), axis.title.y=element_text(size=14)) +
  scale_y_continuous(name="Model Probability") + scale_x_discrete(name="Species")

```

## Now read in the results of the t-test, and zap all the ambiguous datasets
```{r}
tt<-read.table(file="~/Dropbox/Crandall_tobo/scripts/Final_BestModeltable_Using_Means_of_3reps.csv",header=T,sep=",")
#zap the ambiguous datasets
gooddata<-tt$dataset[tt$p.value.Best.Mean...Second.Mean<0.05]
goodmeans<-means[means$dataset %in% gooddata,]

goodprobs<-ggplot(data=goodmeans,mapping=aes(x=dataset,y=modelprob,fill=model)) + 
  geom_bar(stat="identity") + scale_fill_manual(values=my.cols, guide = guide_legend(title = "Migrate Models")) + theme(axis.text.x=element_blank(), legend.text = element_text(size = 14),
                 legend.title = element_text(size = 20), axis.text.y= element_text(size = 14), axis.title.y=element_text(size=14)) + scale_y_continuous(name="Model Probability")
```

## Some Venn diagram action
```{r venn}
library(VennDiagram)
#lump the models into 4 groups
lump<-list(SteppingStone=c("k <= 14, Stepping-Stone, ~40 parameter","k <= 14, Stepping-Stone, 2 parameter"), Regional=c("k = 3, 2 Currents, 7 parameter","k = 2, 1 Current, 4 parameter","k = 2, High/Low Islands, 4 parameter","k = ? Empirical Structure, ? parameter"), Island="k <= 14, N-Island, 2 parameter", Panmixia="k=1, Panmixia, 1 parameter")
means$model2<-means$model
levels(means$model2)<-lump

y<-table(means$dataset,means$model2,means$modelprob>0.01)[,,1]

Stepping.Stone<-means$dataset[which(means$model2=="SteppingStone" & means$modelprob>0.05)]
Regional<-means$dataset[which(means$model2=="Regional" & means$modelprob>0.05)]
Island<-means$dataset[which(means$model2=="Island" & means$modelprob>0.05)]
Panmixia<-means$dataset[which(means$model2=="Panmixia" & means$modelprob>0.05)]
Stepping.Stone<-as.character(Stepping.Stone)
Regional<-as.character(Regional)
Regional<-Regional[-12] #remove mulfla
Island<-as.character(Island)
Panmixia<-as.character(Panmixia)
Chaos<-"mulfla_CB"
x<-list(Stepping.Stone=Stepping.Stone,Regional=Regional,Island=Island,Panmixia=Panmixia)
venn.diagram(x,file="test.tiff")
```

```{r kvalues}
k<-read.table("~/Dropbox/Crandall_tobo/K-values_final.txt",sep="\t",header=T,na.strings = "NA",stringsAsFactors = F)

legtheme<-theme(axis.text = element_text(size = 20),
                 axis.title = element_text(size = 16),
                 legend.text = element_text(size = 18),
                 legend.title = element_text(size = 18),
                axis.title.x = element_blank(),axis.title.y=element_blank())

gkselkoe<-ggplot(data=k) + geom_histogram(mapping=aes(x=Selkoe), binwidth=1, fill="white",color="blue") + ylim(0,14) + scale_x_continuous(limits = c(0,14), breaks=seq(0,14, by=2)) + legtheme

gkmigrate<-ggplot(data=k) + geom_histogram(mapping=aes(x=Migrate), binwidth=1, fill="white",color="blue") + ylim(0,14) + scale_x_continuous(limits = c(0,14), seq(0,14, by=2)) + legtheme

```
